---
title: "Capstone Project"
author: "Praneeth"
date: "October 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install Libraries

```{r}
library(dplyr)
library(tidyverse)
```

## Input Data

```{r}
source("functions_and_data.R")

calls <- readRDS("data/rdas/calls.rda")
call_incdnt <- readRDS("data/rdas/call_incdnt.rda")
incdnts <- readRDS("data/rdas/incdnts.rda")
call_actv <- readRDS("data/rdas/call_actv.rda")
call_actv_typ <- readRDS("data/rdas/call_actv_typ.rda")
call_assoc <- readRDS("data/rdas/call_assoc.rda")
sls_lines <- readRDS("data/rdas/sls_lines.rda")
wrk_grp <- readRDS("data/rdas/wrk_grp.rda")

```

## Join Datasets & EDA

```{r}
#09/26

View(calls)
summary(calls$call_typ_cd)
str(calls$call_typ_cd)
calls_Inbounce <- subset(calls, call_typ_cd == "Inbound ")
View(calls_Inbounce)
summary(calls$call_start_dttm)
Date<- format(as.POSIXct(strptime(calls_Inbounce$call_start_dttm,"%Y-%m-%d %H:%M:%S",tz="")) ,format = "%Y-%m-%d")
Hour<- format(as.POSIXct(strptime(calls_Inbounce$call_start_dttm,"%Y-%m-%d %H:%M:%S",tz="")) ,format = "%Y-%m-%d %H")
calls_Inbounce$ds <- Date
calls_Inbounce$hour <- Hour
Hour<-format(as.POSIXct(strptime(calls_Inbounce$hour,"%Y-%m-%d %H",tz="")) ,format = "%Y-%m-%d %H:%M:%S")
calls_Inbounce$hour <- Hour
View(calls_Inbounce)
ts <- calls_Inbounce %>%
  group_by(Date) %>%
  summarise(y = n())
ts

ts2014 <- subset(ts, ds <= as.POSIXct('2014-12-31  '))
ts2015 <- subset(ts, ds > as.POSIXct('2014-12-31 '))

# prophet
m <- prophet(ts, seasonality.mode = 'multiplicative',
             changepoint.prior.scale = 0.01)
future <- make_future_dataframe(m, periods = 365)
tail(future)
forecast <- predict(m, future)
tail(forecast[c('ds', 'yhat', 'yhat_lower', 'yhat_upper')])
plot(m, forecast) + add_changepoints_to_plot(m)
prophet_plot_components(m, forecast)
df.cv <- cross_validation(m, initial = 180, period = 45, horizon = 90, units = 'days')
head(df.cv)
df.p <- performance_metrics(df.cv)
head(df.p)
plot_cross_validation_metric(df.cv, metric = 'mape')

#The number of call by day, month
Date<- format(as.POSIXct(strptime(calls$call_start_dttm,"%Y-%m-%d %H:%M:%S",tz="")) ,format = "%Y-%m-%d")
Month<- format(as.POSIXct(strptime(calls$call_start_dttm,"%Y-%m-%d %H:%M:%S",tz="")) ,format = "%Y-%m")
calls$date <- Date
calls$month<- Month

call_day <- calls %>%
  group_by(date) %>%
  summarise(y = n())
# Calls trend by date
ggplot(call_day, aes(call_day$date,call_day$y))+
  geom_point()

call_month <- calls %>%
  group_by(month) %>%
  summarise(y = n())
call_month
# Calls by month not working
ggplot(call_month, aes(call_month$month,call_month$y))+
  geom_line()

# number of calls per incident
call_incident_combined <-left_join(call_incdnt, calls, by = "call_id")
call_per_incident <- call_incident_combined %>%
  group_by(incdnt_id) %>%
  summarise(number_of_calls = n())
summary(incident_per_call) # Mean = 1.201

# number of incidents per call
incident_per_call <- call_incident_combined %>%
  group_by(call_id) %>%
  summarise(number_of_incident = n())
summary(incident_per_call) # Mean = 1.509

```

## Code from Daniel

```{r}
calls <- readRDS("data/rdas/calls.rda")

daily_ib_calls_2014 <- calls %>% 
  filter(year(call_start_dttm) == 2014) %>% 
  count(dt = floor_date(call_start_dttm, "1 day"))

daily_ib_calls %>% 
  ggplot(aes(dt, n)) +
  geom_line()

daily_ib_calls %>% 
  mutate(weekend_ind = if_else(wday(dt) %in% c(7,1), "Weekend", "Weekday")) %>% 
  ggplot(aes(dt, n, color = weekend_ind)) +
  geom_line()

#Create a dataset of daily calls for 2014 and 2015 and align on day of week
yoy <- calls %>% 
  transmute(yr = year(call_start_dttm),
            dt = floor_date(call_start_dttm, "1 day"),
            dt = if_else(yr == 2014, dt + years(1) - days(1), dt) %>% ymd
  ) %>% 
  count(yr, dt) %>% 
  spread(yr, n) %>% 
  mutate(err = `2015` - `2014`,
         abs_err = abs(err),
         abs_perc_err = round(abs_err / `2015`, 3) * 100
  ) %>% 
  select(-`2014`, -`2015`) %>% 
  gather(key, value, -dt)

#Compare ME, MAE, and MAPE
yoy %>% 
  ggplot(aes(dt, value, color = key)) +
  geom_line() +
  facet_grid(key~., scales = "free_y")

#Identify days with abnormally large perc error - holidays?
yoy %>% 
  filter(key == "abs_perc_err") %>% 
  mutate(label = ifelse(value >= 40, dt, NA)) %>% 
  ggplot(aes(dt, 
             value, 
             label = label %>% 
               as.Date(origin = "1970-01-01") %>% 
               format("%b-%d")
  )
  ) +
  geom_line() +
  geom_text(color = "red") +
  coord_cartesian(ylim = c(0,100)) +
  scale_y_continuous(expand = c(0,0),
                     breaks = seq(0, 100, 20),
                     minor_breaks = seq(0, 100, 5)
  )

# If we exclude a few extraordinary days where holidays are misaligned, average error is under 800 calls/day if we only use 2014 inbound call count to predict 2015 call count and nothing else.  Holidays can obviously be adjusted and a trend can be applied.

yoy %>% 
  filter(key == "abs_err") %>% 
  ggplot(aes(dt, 
             value
  )
  ) +
  geom_line() +
  geom_hline(yintercept = yoy %>% 
               filter(key == "abs_err",
                      value < 3000
               ) %>% 
               {mean(.$value)},
             color = "red",
             linetype = "dashed"
  ) +
  scale_y_continuous("Number of Calls (Error)",
                     expand = c(0,0),
                     breaks = seq(0,12000, 1000),
                     minor_breaks = seq(0,12000, 500)
  ) +
  scale_x_date(element_blank(),
               date_minor_breaks = "1 month"
  )
```

## Naive Model
```{r mean model building}

#mean model by year - no data split
calls2014 <- subset(calls, call_start_dttm <= as.POSIXct('2014-12-31 23:00'))
calls2015 <- subset(calls, call_start_dttm >= as.POSIXct('2014-12-31 23:00'))

nrow(calls2014)
# 2582385/365 = 7075.027
nrow(calls2015) #2810484/365 = 7699.956

Date<- format(as.POSIXct(strptime(calls2015$call_start_dttm,"%Y-%m-%d %H:%M:%S",tz="")) ,format = "%Y-%m-%d")

calls2015$date <- Date

calls2015num <- calls2015 %>%
 group_by(calls2015$date) %>%
 summarise(y = n())

calls2015num$predicted <- 7075.027 # make new column 
head(calls2015num)

calls2015num$change <- abs(calls2015num$y - calls2015num$predicted)

mean(calls2015num$change) #2726.725 Mean absolute error


#take test and train dataset mean model with data split

set.seed(100)
rows <- createDataPartition(calls2014$call_id, p = .5, list = FALSE)
train <- calls2014[rows, ]
test <- calls2014[-rows, ]

Date<- format(as.POSIXct(strptime(train$call_start_dttm,"%Y-%m-%d %H:%M:%S",tz="")) ,format = "%Y-%m-%d")

train$date <- Date

trainnum <- train %>%
 group_by(train$date) %>%
 summarise(y = n())

View(trainnum)
mean(trainnum$y)#3527.85 because this figure represents 50% of the data, we take 3527.85/.5 = 7055.70 to approximate the rest of the data. 

#compare 2014 train to test
Date<- format(as.POSIXct(strptime(test$call_start_dttm,"%Y-%m-%d %H:%M:%S",tz="")) ,format = "%Y-%m-%d")

test$date <- Date

testnum <- test %>%
 group_by(test$date) %>%
 summarise(y = n())
testnum$predicted <- 3527.85 # make new column 
head(testnum)
testnum$change <- abs(testnum$y - testnum$predicted)

mean(testnum$change) #1146.314 Mean absolute error times 2 because this is half the data 2292.628

#compare 2014 train to 2015 whole

Date<- format(as.POSIXct(strptime(calls2015$call_start_dttm,"%Y-%m-%d %H:%M:%S",tz="")) ,format = "%Y-%m-%d")

calls2015$date <- Date

calls2015num <- calls2015 %>%
 group_by(calls2015$date) %>%
 summarise(y = n())
calls2015num$predicted <- 7055.70 # make new column 
head(calls2015num)
calls2015num$change <- abs(calls2015num$y - calls2015num$predicted)

mean(calls2015num$change) #2734.126 Mean absolute error

#Mean Model
#Summary of MAE
#no data split 2014 to predict 2015 2726.725
#data split 50/50 of 2014 one half to predict the other half 2292.628
#data split 50/50 on 2014 train to predict 2015 2734.126

```

## Create Test, Train datasets & build Auto-ARIMA model
```{r}
call_day_2014 <- calls2014 %>%
  group_by(date) %>%
  summarise(y = n())

call_day_2014<- call_day_2014[1:365,]

call_day_2015full <- calls2015 %>%
  group_by(date) %>%
  summarise(y = n())

# Couldn't sort as date is chr, so split the data, add and combine later
# call_day_2015[with(call_day_2015, order(date)),]

# split data
call_day_2015sub1 <- call_day_2015full[1:196,]
call_day_2015sub2 <- call_day_2015full[197:363,]

newrow1<-data.frame(date='2015-07-16', y= (filter(call_day_2014, date == "2014-07-17")$y))
newrow2<-data.frame(date='2015-07-17', y= (filter(call_day_2014, date == "2014-07-18")$y))

call_day_2015 <- rbind(call_day_2015sub1,newrow1)
call_day_2015 <- rbind(call_day_2015,newrow2)
call_day_2015 <- rbind(call_day_2015,call_day_2015sub2)

arima2014 <- call_day_2014[,2]
arimadata2014 <- ts(arima2014)

fit <- auto.arima(arimadata2014, seasonal = TRUE)
summary(fit) #Train MAE = 1942

xyz<-predict(fit, 365)
call_day_2015$predict <- xyz$pred

ArimaMAE <- mean(abs(call_day_2015$y - call_day_2015$predict))
ArimaMAE #Test MAE = 2448

```

# Visualizations - Calls
```{r}
# Calls by day of week
callsbyday <- rbind(call_day_2014,call_day_2015)

callsbyday$weekday <- weekdays(as.Date(callsbyday$date))
callsbyday$weekday <- factor(callsbyday$weekday)
# levels(callsbyday$weekday) <- c('Sunday', 'Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday')

ggplot(callsbyday, aes(x=weekday, y=y)) + stat_summary(fun.y="mean", geom="bar")

#*** Average calls per day of week
ggplot(call_day, aes(x=factor(dayofweek), y=count)) + stat_summary(fun.y="mean", geom="bar") + ggtitle("Average calls per day")
+ labs(y="Number of Calls per day", x = "Day of Week") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))

#*** Average calls per day in month
ggplot(call_day, aes(x=factor(month), y=count)) + stat_summary(fun.y="mean", geom="bar")+ ggtitle("Average Calls per Month") + labs(y="Number of Calls per day", x = "Month") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
```
# Visualizations - Call duration
```{r}
#with assoc_party_id in each call_id
# is average call_duration time by day, month, workgroup and associate party
call_dur <-left_join(calls, call_assoc, by = "call_id")
Date<- format(as.POSIXct(strptime(call_dur$call_start_dttm,"%Y-%m-%d %H:%M:%S",tz="")) ,format = "%Y-%m-%d")
Month<- format(as.POSIXct(strptime(call_dur$call_start_dttm,"%Y-%m-%d %H:%M:%S",tz="")) ,format = "%Y-%m")
Month<- format(as.POSIXct(strptime(Month,"%Y-%m",tz="")) ,format = "%Y-%m-%d")
call_dur$date <- Date
call_dur$month <- Month
summary(call_dur)

call_dur_day <- call_dur %>%
  group_by(date) %>%
  summarise_at(vars(assoc_durtn_tm), funs(mean(., na.rm=TRUE)))

call_dur_asso <- call_dur %>%
  group_by(assoc_party_id) %>%
  summarise_at(vars(assoc_durtn_tm), funs(mean(., na.rm=TRUE)))

call_dur_month <- call_dur %>%
  group_by(month) %>%
  summarise_at(vars(assoc_durtn_tm), funs(mean(., na.rm=TRUE)))

call_dur_workgroup <- call_dur %>%
  group_by(wrk_grp_id) %>%
  summarise_at(vars(assoc_durtn_tm), funs(mean(., na.rm=TRUE)))
  
# Add date Indicators
call_dur_day$dayofweek <- weekdays(as.Date(call_dur_day$date))
call_dur_day$month <- months(as.Date(call_dur_day$date))
call_dur_day$date <-as.Date(call_dur_day$date)
str(call_dur_day)
call_dur_day$holiday <- ifelse(call_dur_day$date %in% holiday1415, 1, 0)

# linechart and ordered (month)
ggplot(call_dur_day, aes(x=factor(month), y=assoc_durtn_tm)) +
  stat_summary(fun.y="mean", geom="bar") +
  ggtitle("Average calls per day") +
  labs(y="Number of Calls per day", x = "Day of Week")

ggplot(call_dur_day, aes(x=factor(holiday), y=assoc_durtn_tm)) +
  stat_summary(fun.y="mean", geom="bar") +
  ggtitle("Average calls per day") +
  labs(y="Number of Calls per day", x = "Day of Week")

# Call duration by work group names
call_dur_wrkgrp <- left_join(call_dur_workgroup, wrk_grp, by = "wrk_grp_id")
ggplot(call_dur_wrkgrp, aes(x=factor(wrk_grp_desc), y=assoc_durtn_tm)) +
  stat_summary(fun.y="mean", geom="bar") +
  ggtitle("Average calls per day") +
  labs(y="Number of Calls per day", x = "Day of Week")

# no. of calls by wrk grp
#calls_workgroup <- call_dur %>%
  group_by(wrk_grp_id) %>%
  count(unique(call_id))
  
